How to we handle the categorical predictors?

### 2021-11-17 understanding_mm_special.rmd
Population-based...

```{r depend_eff_pred}
depend_eff_pred <- combinepreds(depend_eff_mod, c("emmeans", "varpred", "Effect"), c("gender","rel")
	, x.var = "gender"
	, weights="proportional"
	, x.var.factor = TRUE, plotit = TRUE
)
print(depend_eff_pred 
	+ facet_wrap(~rel, scales = "free_y")
	+ scale_colour_manual(values=cols)
)
```

@@

```{r unbalanced_sim, echo=TRUE}
N <- 100
quant <- seq(0,1,length.out=50)
## Betas
b0 <- 10
b_age <- 0.8
b_jew <- 1
b_muslim <- 0.4
b_male <- 2
b_male_jew <- 0.7
b_male_muslim <- 0.2
betas <- c(b0, b_age, b_jew, b_muslim, b_male, b_male_jew, b_male_muslim)

@@

@@

- CIs excluded for clarity

```{r unbalanced_age}
unbalanced_age <- combinepreds(unbalanced_mod, c("emmeans", "varpred", "Effect"), "age"
	, x.var = "age"
	, at=list(age=age_at)
	, xlevel=list(age=age_at)
	, nesting=NULL
	, weights="proportional"
	, plotit = TRUE
	, ci=FALSE
)
cat("Truth: \n", mean(unbalanced_df$hhsize))
print(unbalanced_age
	+ scale_colour_manual(values=cols)
)
```

>>>>

```{r depend_eff_pred}
depend_eff_pred <- combinepreds(depend_eff_mod, c("emmeans", "varpred", "Effect"), c("gender","rel")
	, x.var = "gender"
	, weights="proportional"
	, x.var.factor = TRUE, plotit = TRUE
)
print(depend_eff_pred 
	+ facet_wrap(~rel, scales = "free_y")
	+ scale_colour_manual(values=cols)
)

# varpred(depend_eff_mod, "gender")
# varpred(depend_eff_mod, "gender", bias.adjust="population")
```

@@

```{r unbalanced_sim, echo=TRUE}
N <- 1000
quant <- seq(0,1,length.out=100)
## Betas
b0 <- 10
b_age <- 0.8
b_jew <- 1
b_muslim <- 0.4
b_male <- 2
b_male_jew <- 0.7
b_male_muslim <- 0.2
betas <- c(b0, b_age, b_jew, b_muslim, b_male, b_male_jew, b_male_muslim)

@@

We are interested in predicted _hhsize_ at the model center. We can explore this in two different ways:

1. **Center of the focal predictor**
$$\mathrm{E(hhsize|\{\bar{x}_{\{f\}}, \bar{x}_{\{n\}}\})}$$

2. **Averaged over levels of focal predictor**
$$\mathrm{mean ~ E(hhsize_i|\{x_{i\{f\}}, \bar{x}_{\{n\}}\})}$$

A few challenges:

- How do we get the representative samples of $x_{i\{f\}}$?
	- `quantiles`
	- `seq`
	- `entire population`

In our simulation example:
\begin{align}
\mathrm{E(hhsize|age)} &= \mathrm{\beta_0 + \beta_A Age + \beta_J Jew + \beta_M Muslim + \beta_M Male}\nonumber\\
&+ \mathrm{\beta_{MJ} Male \times Jew + \beta_{MM} Male \times Muslim}
\end{align}

The non-focal categorical interactions can be handled in two ways:

- product of interacting variables (**effects**)
- model center, i.e., average of all columns of the variables in model matrix (**varpred** and **emmeans**)

@@

Using Eqn. 2 with approach 1 (center of the focal predictor)

<br>

```{r unbalanced_age}
mean_age <- mean(unbalanced_df$age)
mean_hhsize <- mean(unbalanced_df$hhsize)
unbalanced_age <- combinepreds(unbalanced_mod, c("emmeans", "varpred", "Effect")
	, focal = "age"
	, x.var = "age"
	, at=list(age=mean_age)
	, xlevel=list(age=rep(mean_age,5))
	, nesting=NULL
	, weights="proportional"
	, plotit = FALSE
	, ci=FALSE
)
unbalanced_age <- (unbalanced_age
	%>% group_by(model)
	%>% summarize(fit=mean(fit))
	%>% mutate(age=mean_age
		, dev=abs(fit-mean_hhsize)
	)
	%>% data.frame()
)
unbalanced_age
```

An alternative is using Eqn 2 with approach 2

<br>

```{r unbalanced_age_levels}
mean_age <- mean(age_at)
unbalanced_age_levels <- combinepreds(unbalanced_mod, c("emmeans", "varpred", "Effect")
	, focal = "age"
	, x.var = "age"
	, at=list(age=age_at)
	, xlevel=list(age=age_at)
	, nesting=NULL
	, weights="proportional"
	, plotit = FALSE
	, ci=FALSE
)
unbalanced_age_levels <- (unbalanced_age_levels
	%>% group_by(model)
	%>% summarize(fit=mean(fit))
	%>% mutate(age=mean_age
		, dev=abs(fit-mean_hhsize)
	)
	%>% data.frame()
)
unbalanced_age_levels
```

**The later heavily depends on the choice of the levels of the focal predictor**
